---
title: "identify probable multi-unit building closures"
author: "Devraj Kori"
date: "3/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(sf)
library(leaflet)
library(ggmap)
library(rgdal)
library(tidycensus)
library(ggplot2)
library(dummies)
library(estimatr)
library(readxl)
library(ggrepel)
```

Goal: To identify physical addresses that voucher data patterns suggest were multi-unit buildings that were either condemned/vacated entirely or had voucher recipients driven out. 

To qualify in this analysis, a physical address has to:

- be in East End
- meet a minimum threshold of households in any one year of our study period (currently set to 6)
- have no remaining voucher residents by a certain date (currently 31 December 2018)


```{r parameters}
#selct neighborhoods to include
included_neighborhoods<-c("East Liberty", "Homewood","Homewood South","Homewood North", "Larimer",
                          "Point Breeze", "East Hills", "Swisshelm Park",
                          "Shadyside", "Squirrel Hill", "Garfield", "Bloomfield", "Highland Park",
                          "Greenfield", "Hazelwood","Glen Hazel", "Lincoln-Lemington-Belmar",
                          "Lower Lawrenceville", "Upper Lawrenceville", "Central Lawrenceville",
                          "Morningside", "Stanton Heights")
#define parameters that might change

#define minimum number of units in physical address in year to qualify
household_threshold = 6

#define date by which everyone needs to be moved out to qualify
moved_out_by = as.Date("2018-12-31",format("%Y-%m-%d"))

#define date to impute for null moveouts
#setting to 1920 is equivalent to dropping null moveout dates
to_impute_moveout = as.Date("1920-01-23",format="%Y-%m-%d")

#define the number of unique addresses to plot in each line chart
graph_threshold = 6
```

```{r prep_data, include=FALSE}
setwd("..")

# read in geocoded addresses
load("data preparation/geocode addresses/all recipient addresses geocoded.Rdata",verbose=TRUE)
# read in voucher data
load("data preparation/all data clean 29-feb-2020.RData",verbose=TRUE)

#load pittsburgh neighborhood shapefile
neighborhoods<-read_sf("data preparation/geocode addresses/Neighborhoods_/Neighborhoods_.shx")%>%
  select(hood,geometry)

#join dat_cleaned5 with 
geocoded_data<-dat_cleaned5%>%
  # impute "2020-01-23" for missing move-outs
  replace_na(list(MOVEOUTDATE = to_impute_moveout))%>%
  #create the for_geocode_addr column
  mutate(for_geocode_addr = paste(PRIMARYSTREET,ZIP,sep = ", "))%>%
  left_join(correctly_coded%>%
              select(for_geocode_addr))%>%
  select(HA, CLIENT_ID, GENDER, RACE, MOVEINDATE, MOVEOUTDATE, for_geocode_addr,geometry)%>%
  #make into sf object for join
  st_as_sf(crs = "+init=epsg:4326")%>%
  #join with neighborhoods
  st_join(neighborhoods,join=st_within,left=TRUE)%>%
  unique()%>%
  filter(hood%in%included_neighborhoods)

```
```{r no_res_end,echo=FALSE,warning=FALSE,message=FALSE}
# identify physical addresses that had a resident as of moved_out_by
not_closed<-geocoded_data%>%
  #check if moved_out_by is less than moveout
  filter(moved_out_by<=MOVEOUTDATE)%>%
  select(for_geocode_addr)%>%
  unlist()
  

```

```{r construct_panel_frame, echo=FALSE,warning=FALSE,message=FALSE}
year_range<-2003:2019
for(year in year_range){
  year_frame<-geocoded_data%>%
    #filter out the addresses that have residents at the moved_out_by date
    filter(!for_geocode_addr%in%not_closed)%>%
    #create movein and moveout year variables to filter by
    mutate(movein_year=as.numeric(format(MOVEINDATE,"%Y")),
           moveout_year=as.numeric(format(MOVEOUTDATE,"%Y")))%>%
    #filter to only include rows where the stay includes that year
    filter(movein_year<=year & moveout_year>=year)%>%
    #create a flag indicating if they moved out in that year
    mutate(moved_out=(moveout_year==year))%>%
    #remove move-in/move-out fields since they're not relevant in this structure
    select(-c(MOVEINDATE,MOVEOUTDATE,movein_year,moveout_year))%>%
    #add a year column
    mutate(year=year)
  #bind results
  if(year==year_range[1]){
    panel_frame<-year_frame
  }else{
    panel_frame<-rbind(year_frame,panel_frame)
  }
}

```

```{r hh_year,echo=FALSE,warning=FALSE,message=FALSE}
#calculate the number of unique clients per year per address
hh_year_addr<-panel_frame%>%
  group_by(year,for_geocode_addr)%>%
  summarise(hh=n_distinct(CLIENT_ID))

#identify addresses that at least as many households at the thresholds at one point
met_threshold<-hh_year_addr%>%
  group_by(for_geocode_addr)%>%
  summarise(max_hh=max(hh))%>%
  ungroup()%>%
  filter(max_hh>=household_threshold)%>%
  arrange(desc(max_hh))


closed_down<-hh_year_addr%>%
  filter(for_geocode_addr%in%met_threshold$for_geocode_addr)

#unique(closed_down$for_geocode_addr)

```

```{r plot_res_year,echo=FALSE,warning=FALSE,message=FALSE}
# the amount of graphs to make is the number of addresses that met threshold divided by the graph_threshold
to_graph<-1:floor(nrow(met_threshold)/graph_threshold)


for(i in to_graph){
  begin=(i-1)*graph_threshold+1
  end=i*graph_threshold
  addresses_to_graph<-met_threshold%>%
    slice(begin:end)
  #create a dataframe for labels
label_frame<-closed_down%>%
  filter(for_geocode_addr%in%addresses_to_graph$for_geocode_addr)%>%
  group_by(for_geocode_addr)%>%
  #only apply to max year
  summarise(year=round(max(year)))%>%
  as.data.frame()%>%
  select(-geometry)%>%
  #join closed_down back in to get the hh for that year
  left_join(closed_down%>%
              as.data.frame%>%
              select(-geometry))%>%
  #create a field for label
  mutate(for_lab=gsub(",.*","",for_geocode_addr))

print(closed_down%>%
  #select just rows for the top {graph threshold addresses}
  filter(for_geocode_addr%in%addresses_to_graph$for_geocode_addr)%>%
  ggplot(aes(x=year,y=hh,color=for_geocode_addr))+
  geom_line()+
  geom_label_repel(data=label_frame,aes(label=for_lab))+
  scale_color_viridis_d()+
  theme(legend.position="none")+
  labs(x="year",y="voucher households",
       title=)+
  xlim(c(2003,2020))
  )

}





```


```{r plot_leaflet,echo=FALSE}
#read in the bnps
bnps<-read_excel("Redevelopmentprojects_3.24.20.xlsx",sheet="Combined wgeocode")%>%
  #filter out missing xs
  filter(!is.na(X))%>%
  st_as_sf(coords=c("X","Y"),crs=4326)

#map the closed addresses and bnps
leaflet()%>%
  addProviderTiles(providers$CartoDB.Positron)%>%
  addCircleMarkers(data=closed_down%>%
                     group_by(for_geocode_addr)%>%
                     summarise(year=max(year),
                               max_hh=max(hh))%>%
                     ungroup(),
                   stroke=FALSE,
                   radius=~max_hh/3,
                   color="red",
                   popup=~paste(gsub(",.*","",for_geocode_addr),year))%>%
  addCircleMarkers(data=bnps,
                   stroke=FALSE,
                   radius=7,
                   color="blue",
                   popup=~paste(`Development Projects`,`Date-Started`))

```

```{r test, include=FALSE}
test<-hh_year_addr%>%
  filter(grepl("3[0-9][0-9] n negley",for_geocode_addr,ignore.case=TRUE))

```